---
title: "Visualize jsPsych RT Data"
author: "Rick Gilmore"
date: "`r Sys.time()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This document shows how to import a JSON file with data from the jsPsych simple RT experiment and to visualize the data using R.

# Preliminaries

We will need the `jsonlite` and `tidyverse` packages.

```{r, install-pkgs}
if (!require("jsonlite")) {
  install.packages("jsonlite")
}
if (!require("tidyverse")) {
  install.packages("tidyverse")
}
```

# Import data

The data file is in the root (".") directory of the repository.

```{r}
list.files(path = ".", pattern = "\\.json$")
```

Note that we specify a pattern string to limit the search to files that _end_ with `.json` by using the dollar sign (`$`) character.
The pattern string is a [regular expression](https://en.wikipedia.org/wiki/Regular_expression).
Regular expressions are amazingly powerful strings that control how the computer searches for and filters text.
Rick thinks they are like the spells students at Hogwarts used to learn.
He is in no way a Hermione, but you might want to aspire to that level of excellence.
In regular expressions, characters like the period `.` take on different meanings depending on the context.
Here, we want to *search* for a period in the file name, so we have to put two backslashes `\\` before the period `.` to make that work.

Now, let's import the data file.

```{r}
file_name <- list.files(path = ".", pattern = "\\.json$")
if (!is.null(file_name)) (rt_data <- jsonlite::read_json(file_name))
```

Ok, that worked but it's not a tidy data structure.

```{r}
str(rt_data)
```

The `jsonlite::fromJSON` command does a better job parsing the file.

```{r}
rt_data <- jsonlite::fromJSON(file_name)
str(rt_data)
```

# Clean data

We want all of the `test` trials where there was an `image-keyboard-response` as the trial type.

```{r}
rt_data_filtered <- rt_data %>%
  dplyr::filter(., test_part == "test",
          trial_type == "image-keyboard-response")
```

# Plot

Here's a time series.

```{r plot-rt-by-trial-stim-side}
ggplot(rt_data_filtered) +
  aes(x = trial_index, y = rt, color = stimulus, shape = as.factor(key_press)) +
  geom_point() +
  geom_line()
```

What happened on trial 15? Momentary lapse of some sort.

We could make this prettier, but you get the point.
